---
title: 'Worksheet 5: Text Analysis'
author: 'Dora Impson'
date: 'April 6 - 13'
---

_This is the seventh in a series of worksheets for History 8500 at Clemson University. The goal of these worksheets is simple: practice, practice, practice. The worksheet introduces concepts and techniques and includes prompts for you to practice in this interactive document. When you are finished, you should change the author name (above), commit your changes, push to GitHub, and submit the link to your worksheet repository on Canvas. Don't forget to commit your changes as you go - you should make multiple commits as you work through the worksheet, not just one at the end. Also, remember to update your learning log in the `logs/` folder of your repository as you work through the assignment._

Text analysis is an umbrella for a number of different methodologies. Generally speaking, it involves taking a set (or corpus) of textual sources, turning them into data that a computer can understand, and then running calculations and algorithms using that data. Typically, at its most basic level, that involves the counting of words.

**Text analysis can be broken down into 4 general steps:** 

  1. Acquiring a corpus
  2. Preparing the text or Pre-processing
  3. Choosing an analytical tool 
    * (There are many different tools or methods for text analysis. Take a minute and Google each of these methodologies: tf-idf, topic modeling, sentiment analysis, word vector analysis, n-grams)
  4. Analyzing the results
  
In this worksheet we are focusing on basic text analysis. We'll learn how to load textual data into R, how to prepare it, and then how to analyze it using tf-idf or term-frequency according to inverse document frequency. 

Before doing too much, lets load a few relevant libraries. The last few you will likely need to install.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext) 
library(readtext)
library(widyr)
library(SnowballC)
```


## Acquiring a Corpus

First, lets install the State of the Union package. This package contains text of all the state of the Union addresses from Washington to Trump. Run `install.packages` to install the `sotu` package. 
```{r}
install.packages("sotu")
library(sotu)
```

This package includes both the metadata about these speeches in `sotu_meta` and the texts themselves in `sotu_texts`. Lets first look at the metadata associated with this package. 

```{r}
meta <- as.data.frame(sotu_meta)
head(meta)
```

This package also includes a function that will let us write all of the files to disk. This is crucial but also an unusual step because when conducting text analysis in the real world, you will not have an R package filled with the data. Rather you will have to organize the metadata and load the files yourself. Writing these to the disk allows us to practice that step. 

```{r}
file_paths <- sotu_dir(dir = "sotu_files")
head(file_paths)
```

What this does is create a new directory (sotu_files) and adds each State of the Union address as a text file. Notice each speech is its own .txt file that is comprised of just the text of the speech.

(@) Take a look at the directory in your files pane and open one of the documents. 

Now lets load all these texts into R using the `readtext()` function. First look up the documentation for this function and read about it. 
```{r}
sotu_texts <- readtext(file_paths)
```

Take a look at sotu_texts now. Notice that we have two columns, one filled with the text, and one with a document id. 
```{r}
head(sotu_texts, n = 5)
```

Now our textual data is loaded into R but the textual data and the metadata are in two different data frames. Lets combine them. Note that this isn't the way I would typically recommend doing this but its a quirk of the SOTU data. Typically when I create a metadata spreadsheet for a textual dataset I have a column for the file name which makes joining the textual data and metadata together easier. Here, we'll need to sort the dataset so that is alphabetical and then join the two together.

```{r}
sotu_whole <- 
  sotu_meta %>%  
  arrange(president) %>% # sort metadata
  bind_cols(sotu_texts) %>% # pastes the to "spreadsheets" together horizontally- they have to have the same amount rows
  as_tibble() # convert to tibble for better screen viewing
## dhi note to self- notice doc.id is same on sotu_texts and sotu_whole. Also note- this is not the ideal way to do this- should use left_join() with doc.id
glimpse(sotu_whole)
``` 

Now our data is loaded into R and its ready to be pre-processed. 

## Pre-Processing 

### Tokenizing

One of the most basic pre-processing techniques for textual data is to tokenize it. Tokenization is essentially splitting a phrase, sentence, paragraph, or an entire text document into smaller units, such as individual words or terms. Each of these smaller units are called tokens. The tokens could be words, numbers or punctuation marks but, for historians, its common to remove the numbers and punctuation too. To do this we'll create a data frame where each row contains a single word with its metadata as unit of observation.

`tidytext` provides a function called `unnest_tokens().` We can use this to convert our sotu_whole data frame into one that is tokenized. It takes three arguments:
    
    * a tibble or data frame which contains the text
    * the name of the newly created column that will contain the tokens
    * the name of the column within the data frame which contains the text to be tokenized

```{r}
tidy_sotu <- sotu_whole %>%
  unnest_tokens(word, text)

tidy_sotu
```

`unnest_tokens()` also did something else that is really important: it made everything lowercase and took out all punctuation. The function contains options if we wanted to keep those elements, but for our purposes we don't. 

The function `unnest_tokens()` also has an option called token. Tokenizing by word is the default but you could also tokenize by characters, ngrams, lines, or sentences. 

(@)Use the documentation to tokenize the dataset into sentences: 
```{r}
tidy_sotu_sentences <- tidy_sotu %>%
  unnest_tokens(output = "sentence", input = "word", token = "sentences")

tidy_sotu_sentences
```

We've talked about n-grams loosely in class. But lets define it more formally. An n-gram is a contiguous sequence of n items from a given sample of text or speech. The n stands for the number of items. So for example, a bi-gram is sets of two words. 

For example, if I had the string: "Nothing to fear but fear itself" A bi-gram would look like this: 
  Nothing to, to fear, fear but, but fear, fear itself.

A tri-gram would look like this: 
  Nothing to fear, to fear but, but fear itself
  
N-grams are useful for historians when trying to identify specific phrases or terms. For example, if we had a corpus about the 1930s it might be useful to look at bigrams. Thats because analyzing "New Deal" as a bigram would be more effective. If we used unigrams we'd get "new" and "deal" which lose meaning when looked at separately. N-grams let us capture phrases and concepts, not just individual words.

We can use unnest_tokens() to create n-grams for us. To do that we just have to add an extra option that defines n. 
```{r}
sotu_bigrams <- sotu_whole %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2)

head(sotu_bigrams$bigram)
```

(@) Use `unnest_tokens()` to create tri-grams. 
```{r}
sotu_trigrams <- sotu_whole %>%
  unnest_tokens(trigram, text, token = "ngrams", n = 3)

head(sotu_trigrams$trigram)
```

### Stopwords

Another crucial component of text analysis is removing stopwords. Stopwords are words like "I, he, she, of, the" that are common and don't convey meaning. Because they are highly common they don't tell us anything about the content of the text itself. 

There are stopwords that come with the `tidytext` package. 
```{r}
stop_words
```
This is just one example of stopwords. You can find other lists such as stopwords in other languages or [stopwords designed specifically for the 19th century.](https://www.matthewjockers.net/macroanalysisbook/expanded-stopwords-list/) Its also possible you may want to edit the list of stopwords to include some of your own. For example, if we wanted to add the word, "America" to the stopwords list we could use add_row to do so: 
```{r}
stop_words_custom <- stop_words %>% add_row(word="America", lexicon="NA")
```


For now lets just remove the default stopwords. The easiest way to do that here is to do an anti-join. We join and return all rows from our table of tokens tidy_sotu where there are no matching values in our list of stopwords. 

```{r}
tidy_sotu_words <- tidy_sotu %>% 
  anti_join(stop_words)

tidy_sotu_words

#another way to do this would be to filter by words NOT in the stop word list like this:  filter(!word %in% stop_words$word)
```

### Stemming 

The third common kind of pre-process is called word stemming. This process reduces a word to its root stem. So for example: fishing becomes fish, fished becomes fish, fishes becomes fish. You can easily see how this might be useful for capturing all forms of a word.

`tidytext` doesn't have its own word stemming function. Instead we have to rely on the functions provided by `hunspell` or `SnowballC`. I prefer `SnowballC`. You may need to install it before running the below code. 

```{r}
library(SnowballC)
tidy_sotu_words %>%
        mutate(word_stem = wordStem(word))
```

Now if you compare the word and word_stem columns you can see the effect that wordStem had. Notice that it works well in cases like 
  
  citizens = citizen 

But it does some odd things to words like representatives. Whether this is useful for you will depend on the question your asking (and the OCR accuracy) but its a useful technique to be familiar with nevertheless. 

While stemming can be useful for some applications, it's often problematic for historical research. Stemming algorithms strip words down to roots without understanding historical context or meaning. For example, "suffragist" becomes "suffrag," "temperance" becomes "temper," and "reconstruction" becomes "reconstruct"—losing the specific historical concepts these words represent. Even worse, stemming can collapse historically distinct terms into the same stem: "miner" and "minor" both become "min," or "united" and "union" both stem to "unit." For historians, these distinctions matter. A text analysis of Civil War documents that conflates "union" (the political entity) with "united" (an adjective) produces meaningless results. Similarly, analyzing women's reform movements while collapsing "suffragist," "suffragette," and "suffrage" into "suffrag" erases the meaningful differences between these terms. Unless you have a specific reason to use stemming—like dealing with extremely messy OCR or wanting to capture all grammatical forms of a common word—it's often better to skip this step and work with the original words. 

## Analysis

Lets reset our work space and ensure that our df is loaded with single tokenized words and filter by our stopword list. Go ahead and clear your environment by running `rm(list=ls())` in the console. Then run the below code. This code is simply everything we've run up to this point. 

```{r}
meta <- as.data.frame(sotu_meta)
file_paths <- sotu_dir(dir = "sotu_files")
sotu_texts <- readtext(file_paths)
sotu_whole <- 
  sotu_meta %>%  
  arrange(president) %>% # sort metadata
  bind_cols(sotu_texts) %>% # combine with texts
  as_tibble() 

tidy_sotu <- sotu_whole %>%
  unnest_tokens(word, text) %>% 
  anti_join(stop_words)
```
(@) Before we move forward, take a minute a describe the chunk of code you just ran. What does each section do and how does it reflect the workflow for a topic modeling project? What are the important steps that are unique to topic modeling? 

>meta <- as.data.frame(sotu_meta)##this created a new variable that presents the sotu_meta data as as dataframe. Doing so allows for tidy orginalization of data into rows and columns with one entity in each cell.
file_paths <- sotu_dir(dir = "sotu_files")## this step creates a directory called sotu_dir. The creation of a directory provides the framework for the organization and tidying of data.
sotu_texts <- readtext(file_paths) ## this step loads the text files created in the previous step into R and provides two columns, one with the text and one with an accompaning doc_id. The doc_id will allow for future joining of datasets if needed. 
sotu_whole <-  
  sotu_meta %>%  
  arrange(president) %>% # sort metadata
  bind_cols(sotu_texts) %>% # combine with texts
  as_tibble()
  ## these steps create a new variable from the sotu_meta dataframe with has been sorted by president and then joined with the sotu_texts dataframe. They have been joined by doc_id. The new variable which is called sotu_whole is then presented as a tibble which provides an easy way to check and make sure there are not mistakes in the code.
   

tidy_sotu <- sotu_whole %>%
  unnest_tokens(word, text) %>% 
  anti_join(stop_words)
  ## A new variable called tidy_sotu is created from the sotu_whole dataframe which has been passed to the unnest_token function and sorted by word and text columns. This information is then passed onto the anti_join function which applies the stop_word list to the corpus. This entire block of code yields a dataset the is tidy, organized by doc_id and rid of unwanted items (punctuation and stop words). The tidy_sotu dataframe is now ready to apply topic modeling.


The most basic kind of analysis we might be interested in doing is counting words. We can do that easily using the `count()` function: 
```{r}
tidy_sotu %>%
  count(word, sort = TRUE)
```

Now we know that the most used word in state of the union speeches is government. But what if we wanted to look at when presidents use the words war versus the word peace? 
```{r}
tidy_sotu %>%
  filter(word %in% c("war", "peace")) %>% 
  count(year, word)
```

This data frame is to big to understand quickly without visualizing it. We can create a bar chart to better understand it: 
```{r}
library(ggplot2)

tidy_sotu %>%
  filter(word %in% c("war", "peace")) %>% 
  count(year, word) %>% 
  ggplot(aes(year, n, fill = word)) +
    geom_col(position = "fill")

```

We also might want to ask about the average length of each president's state of the union address. Who had the longest speech and who had the shortest?
```{r}
sotu.length <- tidy_sotu %>%
  count(president, doc_id)  %>% 
  group_by(president) %>% 
  summarize(avg_words = mean(n)) %>% 
  arrange(desc(avg_words))
```

(@) Think back to the metadata that we loaded about these speeches. Why are the STOU addresses for more modern president's shorter than others? (For example Bush vs Teddy Roosevelt.) What about the presidency, and the State of the Union tradition, changed that would account for this? 

> The advent of television and internet has most likely changed the modern president's length of speech for the State of the Union address. The speech is now presented on primetime television and there are expectations that it will fill a certain timeslot in programming. 

(@) Filter the dataset to address this discrepancy and the recreate these statistics: 
```{r}
sotu.length.radio <- tidy_sotu %>%
  filter(year >= 1923 & year <= 1946) %>%
  count(president, doc_id) %>% 
  group_by(president) %>% 
  summarize(avg_words = mean(n)) %>% 
  arrange(desc(avg_words))

  sotu.length.radio %>% print(n = Inf)
  ## based on this, I cannot necessarily argue that radio made a huge difference in the length of the SOTU speeches. There does not seem to be a growing pattern as radios became more popular. Next we will look a TV.
 
 sotu.length.television <- tidy_sotu %>%
  filter(year >= 1947 & year <= 2021) %>%
  count(president, doc_id) %>% 
  group_by(president) %>% 
  summarize(avg_words = mean(n)) %>% 
  arrange(desc(avg_words))

  sotu.length.television %>% print(n = Inf)
```
## comparing the lengths of early modern presidents' SOTU  to those with radio and television capablities shows that on the whole, the advent of technology seemed to shorten the average length of SOTU speeches. The limitations placed upon airtime and the addition most likely led to shorter speeches.

### Term Frequency
Often, the raw frequency of a term is not as useful as relative frequency. In other words, how often that word appears relative to the total number of words in a text. This ratio is called **term frequency**. 

You can calculate the term frequency by dividing the total occurrences of a word by the total number of words. Typically you want to do this per document.

Here's an easy way to calculate it: 
```{r}
tidy_sotu_rel.freq <- tidy_sotu %>%
  count(doc_id, word, sort = T)  %>%# count occurrence of word and sort descending
  group_by(doc_id) %>% 
  mutate(n_tot = sum(n), # count total number of words per doc
         term_freq = n/n_tot)
```

We can assume that words with a high frequency in the text are more important or significant. Here we can find the words with the most significance for each president: 
```{r}
tidy_sotu %>%
  count(president, word)  %>%  # count n for each word
  group_by(president) %>% 
  mutate(n_tot = sum(n), # count total number of words per doc
         term_freq = n/n_tot) %>% 
  arrange(desc(term_freq)) %>% # sort by term frequency
  top_n(1) %>%  # take the top for each president
  print(n = Inf) # print all rows
```
(@) The code above is commented to help you follow it. Walk through the code above, and explain what each line does in your own words. If its a function you are unfamiliar with, look up the documentation.

> tidy_sotu %>% ## take the tidy_sotu dataset and pass it to count()
  count(president, word)  %>%  ## This line provides a count for each word broken down by president. So, all the times that "senate" appears will be grouped by President (jefferson, lincoln etc.) It will then pass this info via the pipe operator to the next line.

  group_by(president) %>% ## groups all the info recieved by president. Passes the information on to the next line via pipe

  mutate(n_tot = sum(n),
    term_freq = n/n_tot) %>% 
    
    ## takes the info from the group_by() and makes a new column in the dataset which gives the total sum of words by president. This also creates a function for the frequency terms appear and sets the frequency with the ratio of n (count of words) divided by total number of words. 

          
  arrange(desc(term_freq)) %>% 
  ## sorts the infomation from the new n_tot column by term frequency. Passes this data onto next step via pipe.

  top_n(1) %>%  # take the top for each president
  print(n = Inf) # print all rows

### TF-IDF

The above measures the frequency of terms within individual documents. But what if we know about words that seem more important based on the contents of the **entire** corpus? That is where tf-idf or term-frequency according to inverse document frequency comes in. 

Tf-idf measures how important a word is within a corpus by scaling term frequency per document according to the inverse of the term’s document frequency (number of documents within the corpus in which the term appears divided by the number of documents). The tf–idf value increases proportionally to the number of times a word appears in the document and is offset by the number of documents in the corpus that contain the word, which helps to adjust for the fact that some words appear more frequently in general. 

Don't worry too much about how tf-idf is calculated. But if you feel like you are a bit lost and want to understand the specifics - I recommend reading the [tf-idf wikipedia page](https://en.wikipedia.org/wiki/Tf%E2%80%93idf) and this blog post from [_Learn Data Science_](https://www.learndatasci.com/glossary/tf-idf-term-frequency-inverse-document-frequency/).

We'll calculate tf-idf in the next code chunk but lets talk for a second about what that number will represent. It will be: 

    * lower for words that appear frequently in many documents of the corpus, and lowest when the word occurs in virtually all documents.
    * higher for words that appear frequently in just a few documents of the corpus, this lending high discriminatory power to those few documents.

Luckily, `tidytext` provides a function for calculating tf-idf. To calculate tf-idf the function needs a list of every word in every document and the count. Like this: 
```{r}
tidy_sotu %>%
  count(doc_id, word, sort = TRUE)  
```
We can feed that to the function and get the tf-idf: 
```{r}
sotu.tf.idf <- tidy_sotu %>%
  count(doc_id, word, sort = TRUE)  %>%  
  bind_tf_idf(word, doc_id, n) 

head(sotu.tf.idf)
```

The resulting data frame has 3 columns: term frequency (tf), inverse document frequency (idf) and Tf-idf (tf_idf).

Lets look at what the words with the highest tf-idf score are. Note that common words across all speeches -- things like "government" and "people" -- will have low tf-idf scores even if they are frequent. Thats the point!
```{r}
sotu.tf.idf %>% arrange(desc(tf_idf))
```

(@) Pick a president who served more than one term. Filter the dataset and generate both raw word counts and tf-idf scores for that president. What words are most significant in each method? Why and what does that tell you about that president and their administration?  

```{r}
cleveland_word_counts <- tidy_sotu %>%
  filter(president == "Grover Cleveland") %>%
  count(word, sort = TRUE) ## this gives me the raw word counts for Grover Cleveland.
head(cleveland_word_counts)
## top words were: 1 government   574
2 united       312
3 congress     219
4 people       216
5 public       184
6 service      184
## top raw word scores reflect Cleveland's desire to work on civil service reform and his use of vetos.

cleveland_tfidf <- tidy_sotu %>%
  filter(president == "Grover Cleveland") %>%
  count(doc_id, word, sort = TRUE) %>%
  bind_tf_idf(word, doc_id, n)
  ## gives the tdidf for Cleveland's speeches.

cleveland_tfidf %>%
  arrange(desc(tf_idf))
## this gives the highest tf-idf scores for Cleveland's speeches
head(cleveland_tfidf)
```
## top tf-idf words were government and gold. Knowing that Cleveland believed in fiscal conservatism, limited government and fought to maintain the gold standard these results are not surprising.

(@) What does a high tf-idf score tell us historically? Think about what it means when a word has a high score versus a low score. Why might a word like "watergate" have a very high tf-idf score in Nixon's speeches but a word like "government" has a low score even though it appears frequently? How might you apply this in your own research? What words would you expect to see with high and low scores?

>## High tf-idf scores indicate that a word is unique and distinctive to the dataset and should prompt further inquiry. Low tf-idf scores indicate that a word is common throughout the document. 

### Co-Occurance
Co-occurrence gives us a sense of words that appear in the same text, but not necessarily next to each other. It shows words that are likely to co-occur. Note that this is different than topic modeling, which we'll discuss next week. 

For this section we will make use of the `widyr` package. The function which helps us do this is the `pairwise_count()` function. It lets us count common pairs of words co-appearing within the same speech. This function might take a second as the resulting data frame will be incredibly large. Also note that we're using the last 500 words of each speech because SOTU addresses vary greatly in length. This gives us a consistent chunk size to analyze.

```{r}
sotu_word_pairs <- sotu_whole %>% 
  mutate(speech_end = word(text, -500, end = -1)) %>%  # extract last 500 words
  unnest_tokens(word, speech_end) %>%   # tokenize
  filter(!word %in% stop_words$word) %>%  # remove stopwords
  pairwise_count(word, doc_id, sort = TRUE, upper = FALSE) # don't include upper triangle of matrix

head(sotu_word_pairs)
```

Now we have a list of words that appear near each other in the text as well as the frequency. Once again this dataset is far too large to look at in a data frame. Instead, we'll create a network graph that shows us the relationships between words for any words that appear more than 200 times. I chose 200 after looking at the above dataset and seeing that the highest count was 239. You want the network graph to be manageable and not too large. 
```{r}
library(igraph)
library(ggraph)

sotu_word_pairs %>% 
  filter(n >= 100) %>%  # only word pairs that occur 200 or more times
  graph_from_data_frame() %>% #convert to graph
  ggraph(layout = "fr") + # place nodes according to the force-directed algorithm of Fruchterman and Reingold
  geom_edge_link(aes(edge_alpha = n, edge_width = n), edge_colour = "tomato") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, 
                 point.padding = unit(0.2, "lines")) +
  theme_void()
``` ## changed the word pair setting above because I was getting a blank graph. Based on the head()view, the word-pair count was set too high.
(@) Create a network graph that shows the relationship between words that appear between 125 and 175 times.
## I adjusted this to 75 to 100 times to fit the word-pair count. 
```{r}
sotu_word_pairs %>% 
  filter(n >= 75 & n <= 100) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n), edge_colour = "tomato") +
  geom_node_point(size = 5) +
  geom_node_text(aes(label = name), repel = TRUE, 
                 point.padding = unit(0.2, "lines")) +
  theme_void()

```

## Analyzing Historical Journals

In the github repository below I have included the text and metadata for a journal called _Mind and Body_ which ran from the 1890s until the late 1930s and chronicled the development of the physical education profession. This profession paralleled many important developments in American history through the Progressive Era, World War I, the 1920s, and finally the Great Depression. This dataset was OCR'd from copies stored in Google Books. Using the metadata provided and the raw text files can you use what you learned above to analyze these texts? What historical conclusions might you be able to draw?

Some questions to explore:
- How does the vocabulary change over time periods (Progressive Era → WWI → 1920s → Depression)?
- What words have high tf-idf scores in specific time periods?
- Are there words that co-occur frequently? What historical events or movements might they reflect?
- How does the journal discuss gender, physical fitness, or professionalism?

Note that when you are loading text from a file the process is slightly different. Above we were loading text from a R package which necessitates a different approach and is not something you would do frequently in practice. The readtext function takes a directory of text files and loads them into a dataframe. We also load a metadata file separately. You will need to join those two dataframes together to make use of the information in the metadata file. 
```{r}
#| eval: false
#zip file of all the .txt files. One for each issue. 
download.file("https://github.com/regan008/8510-TextAnalysisData/raw/refs/heads/main/MindAndBody.zip", "MindAndBody.zip")
unzip("MindAndBody.zip")

# Metadata that includes info about each issue.
metadata <- read.csv("https://raw.githubusercontent.com/regan008/8510-TextAnalysisData/main/mb-metadata.csv")

mb <- readtext(paste(getwd(), "/txt/*.txt", sep=""))
```

(@) Add code chunks below and intersperse text to explain what you are doing and why.
```{r}
colnames(metadata)
colnames(mb) ## Asked copilot how to check the column names in both dataframes to see if I could just join by a common column. Ran this code and realized that they do not match exactly. The metadata set has a Filename column while the mb set has a doc_id column. Then I looked at the actual files and realized that they need to be cleaned before I could do a left_join. The mb files have a "MB"identifier in front of the file name. If I take this off, I can then join the files with left_join.
mb_combined <- mb %>%
  mutate(doc_id = str_remove(doc_id, "^MB_")) %>%  # removes "MB_" from the start of the doc_id files(the ^ means beginnig of string)
  left_join(metadata, by = c("doc_id" = "Filename")) # by = c("doc_id" = "Filename") joins the cleaned doc_id with metadata's Filename column. left_join() keeps all rows from mb and adds matching columns from the metadata dataframe.

head(mb_combined) ##checked my work

tidy_mb <- mb_combined %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) # here I am creating a new tidy dataset where I tokenize the text column into individual words (no lowercase, no punctuation) and then I removed stopwords from the data using anti_join and the tidyverse set of stopwords.
View(tidy_mb) # used this to check my work

#raw word count for progressive era (1895 to 1920)

progressive_era_mb <- tidy_mb %>%
  filter(Year >= 1895 & Year <= 1920) %>%
  count(word, sort = TRUE)
  View(progressive_era_mb) #I noticed that there are numeric numbers in the word count. The highest ranking raw "word" is currently 1. So I am going to remove the numbers because I really want to see the most used raw words.
progressive_era_mb <- tidy_mb %>%
  filter(Year >= 1895 & Year <= 1920) %>%
  filter(!grepl("^[0-9]+$", word)) %>%  # removes words that are only numbers
  count(word, sort = TRUE)
View(progressive_era_mb)
# this showed me only the top raw words. It is interesting that they include "physical", "training" and "digitized" within the top 5 raw word counts. These words remind me of the progressive era's focus on modernization and health.

#Next I am going to get the tf_idf for the progressive era data.

progressive_era_tfidf <- tidy_mb %>%
  filter(Year >= 1895 & Year <= 1920) %>%
  filter(!grepl("^[0-9]+$", word)) %>%  
  count(doc_id, word, sort = TRUE) %>% 
  bind_tf_idf(word, doc_id, n) # this built onto my previous code by giving me a word count per document and then calculated the tf-idf on the word counts grouped by document.
progressive_era_tfidf %>%
  arrange(desc(tf_idf)) #this gives the results sorted by tf-idf score.
#After viewing I realize that there are many words with a zero idf score. These words are in every document in the dataset and I probably should have included them in the stop words list (most noteably "physical") I used copilot to filter them out of the results so I can more easily see the tf-idf scores.

progressive_era_tfidf %>%
  filter(idf > 0) %>%
  arrange(desc(tf_idf))
  #this did not work and I am not sure why.I could modify my stopword list to include "physical" and "left", both of which show up frequently and do not add value to the overall tf-idf score.
```





(@) Choose one of the datasets from this repository or use one of your own: https://github.com/regan008/8510-TextAnalysisData. Note that there are both zip files with .txt files as well as separate metadata files for each publication in this repo. 

Write a text analysis script that explores and asks questions of the texts. At the end write a summary that analyzes the results you found. Be sure to rely on your historical knowledge. What can we learn from applying the strategies you've learned in this worksheet to analyze these documents?

```{r}
download.file("https://github.com/regan008/8510-TextAnalysisData/blob/main/TheAmericanCity.zip?raw=true", "AmCity.zip")
unzip("AmCity.zip")

AmCitymetadata <- read.csv("https://raw.githubusercontent.com/regan008/8510-TextAnalysisData/main/AmCityMetadata.csv")
#First I downloaded my .txt files and metadata files for the AmCity data.

AmCity_data <- readtext(paste(getwd(), "/txt/*.txt", sep=""))
# this loads all the .txt files into a subdirectory and all the results are stored in a variable called AmCity_data

AmCity_combined <- AmCity_data %>%
  mutate(doc_id = str_remove(doc_id, "^MB_")) %>%
  left_join(AmCitymetadata, by = c("doc_id" = "filename"))
# I took the AmCity_data and cleaned it up, removing the "MB" at the beginning of the doc_id listings. Then I joined it with the dataset by left_join. 

tidy_AmCity <- AmCity_combined %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) 
#Created a new tidy dataset where I tokenize the text column into individual words (no lowercase, no punctuation) and then I removed stopwords from the data using anti_join and the tidyverse set of stopwords.

progressive_era_AmCity <- tidy_AmCity %>%
  filter(year >= 1895 & year <= 1920) %>%
  count(word, sort = TRUE)
  head(progressive_era_AmCity)
 

  #I decided to get a raw word count for the progressive era. I thought it may be interesting to see what the top words mentioned in this dataset were compared with the top raw word count from the Mind and Body journals for the same years. After running head()to see my top hits for raw word count, I notice that the same issue with the Mind Body data has occured where there are numerals in the word column. I filter to remove them so I can just see words only.

progressive_era_AmCity <- tidy_AmCity %>%
  filter(year >= 1895 & year <= 1920) %>%
  filter(!grepl("^[0-9]+$", word)) %>%  # removes words that are only numbers
  count(word, sort = TRUE)
head(progressive_era_AmCity)
# some overlap between the two journals occured. There is digitized and training in both as top raw word count hits. School is also listed in the top of the raw word count. This can lead me to assume that the progress era was interested in education, exerise, training of some sort, and technology.
```